---
title: "NIST Atomic-Line Log-Cosine Scan — Final Adversarial Statistical Audit"
subtitle: "Global calibration, frozen holdouts, alternative nulls, and preserved failures"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(fig.align = "center", out.width = "92%")
here_root <- normalizePath(file.path(dirname(knitr::current_input(dir = TRUE)), ".."),
                           mustWork = FALSE)
if (is.na(here_root) || !dir.exists(here_root)) here_root <- normalizePath("..", mustWork = FALSE)
TAB <- file.path(here_root, "tables_r/statistical_audit")
FIG <- file.path(here_root, "figures_r/statistical_audit")
rd <- function(f) {
  p <- file.path(TAB, f)
  if (file.exists(p)) utils::read.csv(p, stringsAsFactors = FALSE, check.names = FALSE) else NULL
}
fig <- function(f) {
  p <- file.path(FIG, f)
  if (file.exists(p)) p else NULL
}
show_tab <- function(df, n = NULL) {
  if (is.null(df)) return(cat("_table not available_"))
  if (!is.null(n)) df <- utils::head(df, n)
  knitr::kable(df)
}
```

# 1. Executive verdict

This report is the **final adversarial statistical audit** of the NIST atomic-line
log-cosine scan. It is deliberately stricter than the development audit:
full-resolution Monte Carlo budgets are passed to every module without hidden
500/1,000-trial caps; the confirmatory frequency is frozen before each holdout;
the complete declared preprocessing multiverse enters multiplicity correction;
alternative null generators are tested; and every failed, mixed, inconclusive,
or not-established claim is retained in a separate machine-readable table.

The central global-evidence table is shown below. `largest_reported_global_p` is
the largest among the primary scan-global p-value, family-max/adjusted
multiplicity p-values, and the worst declared alternative-null p-value. This is a
**conservative diagnostic summary only**, not a newly combined formal p-value.

```{r global-evidence}
show_tab(rd("calibrated_global_evidence.csv"))
```

# 2. Claims — including failures

No result is promoted merely because another analysis passes. In particular,
failed held-out predictive transfer, weak bin/specification stability, an
alternative-null failure, or a non-significant frozen holdout remains failed or
mixed even if the in-sample peak is strong.

```{r final-claims}
show_tab(rd("final_adversarial_claim_matrix.csv"))
```

## 2.1 Non-passing claims retained explicitly

`failed_claims.csv` is generated mechanically from the claim matrix and contains
every verdict other than `pass`.

```{r failed-claims}
show_tab(rd("failed_claims.csv"))
```

# 3. Exact run provenance

The manifest is written **before** this report is rendered, so the report cannot
silently embed a manifest from a previous run.

```{r provenance}
show_tab(rd("final_adversarial_run_manifest.csv"))
show_tab(rd("final_adversarial_module_status.csv"))
show_tab(rd("analysis_registry.csv"), 30)
show_tab(rd("dataset_flow_by_species.csv"))
```

# 4. Primary peak and bootstrap uncertainty

Bootstrap uncertainty measures sampling and peak-selection stability within the
same line list. It is not independent experimental replication.

```{r bootstrap}
show_tab(rd("peak_estimates.csv"))
show_tab(rd("peak_confidence_intervals.csv"))
show_tab(rd("peak_stability.csv"))
```

```{r bootstrap-figure}
p <- fig("fig05_bootstrap_peak_distribution.png")
if (!is.null(p)) knitr::include_graphics(p)
```

# 5. Full-scan look-elsewhere correction

Pointwise and scan-global evidence remain separate. The scan-global statistic is
the maximum deviance improvement over the complete declared frequency grid.
Empirical p-values use `(r + 1)/(B + 1)`; zero exceedances are reported at the
Monte Carlo resolution floor rather than as `p=0`.

```{r global-scan}
show_tab(rd("significance_results.csv"))
```

# 6. Full declared-family multiplicity correction

The corrected family contains the predefined Fe II bin grid, the complete
sigma/degree/bin preprocessing multiverse, observed/Ritz representations, and
neighbouring ion-II controls, with duplicate numerical specifications removed.
The audit reports:

- Bonferroni and Holm FWER control, neither of which requires independence;
- Benjamini-Hochberg FDR as a standard reference;
- Benjamini-Yekutieli FDR for arbitrary dependence;
- an approximate family maximum-statistic calibration.

The family-max simulation couples marginal per-analysis null draws rather than
constructing a fully joint synthetic line-list generator, so its dependence
structure remains an explicit limitation. Holm/Bonferroni provide the
arbitrary-dependence FWER cross-check.

```{r multiplicity}
show_tab(rd("multiple_testing.csv"))
```

```{r multiplicity-figure}
p <- fig("fig12_multiplicity_comparison.png")
if (!is.null(p)) knitr::include_graphics(p)
```

# 7. Adversarial alternative null models

The canonical fixed-baseline fitted-Poisson null is supplemented with four
stress tests:

1. fitted Poisson with Gaussian smoothing re-estimated in every replicate;
2. conditional multinomial counts with total line count fixed;
3. negative-binomial overdispersion with a method-of-moments size estimate;
4. block-permuted Pearson-residual rate perturbations followed by Poisson
   sampling and baseline re-estimation.

The conservative result is the largest scan-global p-value among these declared
nulls. These alternatives are **post-signal adversarial stress tests**, not a
claim of historical preregistration or of a unique physical null. The
block-residual construction is specifically a misspecification diagnostic.

```{r alt-nulls}
show_tab(rd("alternative_null_results.csv"))
```

```{r alt-nulls-figure}
p <- fig("fig17_alternative_nulls.png")
if (!is.null(p)) knitr::include_graphics(p)
```

# 8. Synthetic-null false-positive calibration

Calibration asks whether the complete scan-global decision rule achieves its
nominal type-I error under the canonical fitted-Poisson null. The finite Monte
Carlo interval is reported. Robustness to other generators is assessed
separately in the alternative-null section rather than conflated with this
calibration exercise.

```{r calibration}
show_tab(rd("null_calibration.csv"))
```

```{r calibration-figure}
p <- fig("fig13_null_calibration.png")
if (!is.null(p)) knitr::include_graphics(p)
```

# 9. Frozen blocked holdouts

For every confirmatory holdout, `k` is selected only on the training bins and
then locked. The test block is evaluated at that fixed frequency. Exploratory
test-block rescans are reported separately and cannot rescue a failed locked-k
result. The final claim matrix uses a strict rule: all declared holdout designs
must have positive direction and fixed-k `p <= 0.05` to receive `pass`; partial
success is `mixed`.

```{r holdout}
show_tab(rd("holdout_results.csv"))
```

```{r holdout-figure}
p <- fig("fig09_holdout_validation.png")
if (!is.null(p)) knitr::include_graphics(p)
```

# 10. Preprocessing and specification sensitivity

The audit-declared multiverse varies baseline smoothing, polynomial degree, bin
count, and line-position representation. An 80% reference-region fraction is
used as a descriptive robustness threshold in the final claim matrix; this is an
audit criterion, not a universal physical threshold and not represented as
historical preregistration.

```{r sensitivity}
show_tab(rd("bin_stability_summary.csv"))
show_tab(rd("bin_grid_results.csv"))
show_tab(rd("specification_results.csv"))
show_tab(rd("observed_ritz_replication.csv"))
```

```{r sensitivity-figures}
p <- fig("fig10_specification_curve.png"); if (!is.null(p)) knitr::include_graphics(p)
p <- fig("specification_heatmap.png"); if (!is.null(p)) knitr::include_graphics(p)
```

# 11. In-sample versus held-out model comparison

AIC/BIC improvement after selecting a frequency on the same data does not by
itself demonstrate predictive transfer. Held-out log-likelihood is reported
separately. A negative `M1 - M0` held-out gain is retained as a failed claim even
if the in-sample improvement is very large.

```{r model-comparison}
show_tab(rd("model_comparison.csv"))
```

```{r model-comparison-figure}
p <- fig("fig11_model_comparison.png")
if (!is.null(p)) knitr::include_graphics(p)
```

# 12. Injection and recovery

Injection/recovery reports both detection of any significant peak and recovery
inside the correct frequency region. A detector that fires at the wrong
frequency has not recovered the injected structure. The zero-amplitude row is
also retained as a direct type-I diagnostic.

```{r injection}
show_tab(rd("injection_recovery.csv"))
```

```{r injection-figure}
p <- fig("fig14_injection_recovery_power.png")
if (!is.null(p)) knitr::include_graphics(p)
```

# 13. Computational parity

```{r parity}
show_tab(rd("python_r_parity.csv"))
```

```{r parity-figure}
p <- fig("fig15_python_r_parity.png")
if (!is.null(p)) knitr::include_graphics(p)
```

# 14. Interpretation boundaries

This audit can support or weaken a claim of statistical structure in the
declared NIST analysis. It cannot by itself establish:

- independent experimental confirmation;
- independence of observed/Ritz representations or blocked subsets drawn from
  the same transition list;
- a unique physical null model;
- a WCT mechanism, causal explanation, or universal atomic law;
- NIST endorsement of the analysis.

Computational parity, same-dataset bootstrap stability, frozen holdouts,
alternative-null robustness, and independent physical replication are therefore
treated as different evidentiary categories.

# 15. Reproduction command

The strict final runner fails closed and never substitutes development budgets.

```{r reproduce}
cat("Rscript R/run_final_adversarial_audit.R --parallel true --render-report true\n\n")
manifest <- rd("final_adversarial_run_manifest.csv")
if (!is.null(manifest)) show_tab(manifest)
cat("\nSession information for this render:\n")
sessionInfo()
```
