---
title: "NIST Atomic-Line Log-Cosine Spectral Scan — Statistical Audit"
subtitle: "Reproducible statistical reporting standard for the Fe II analysis suite"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(fig.align = "center", out.width = "92%")
here_root <- normalizePath(file.path(dirname(knitr::current_input(dir = TRUE)), ".."),
                           mustWork = FALSE)
if (is.na(here_root) || !dir.exists(here_root)) here_root <- normalizePath("..", mustWork = FALSE)
TAB <- file.path(here_root, "tables_r/statistical_audit")
FIG <- file.path(here_root, "figures_r/statistical_audit")
rd <- function(f) { p <- file.path(TAB, f); if (file.exists(p)) utils::read.csv(p, check.names = FALSE) else NULL }
fig <- function(f) { p <- file.path(FIG, f); if (file.exists(p)) p else NULL }
show_tab <- function(df, n = NULL) {
  if (is.null(df)) return(cat("_table not available_"))
  if (!is.null(n)) df <- utils::head(df, n)
  knitr::kable(df)
}
```

# 1. Executive summary

This report is a **statistical audit** of an independent R computational
reproduction of the canonical Python NIST log-cosine spectral scanner. It does
**not** claim independent experimental validation or any physical mechanism.

**Supported by this audit (computational / statistical):**

- deterministic R/Python computational agreement on the canonical pipeline;
- the observed Fe II scan results and scan-global bootstrap significance;
- measured bin stability across the predefined 60–240 bin grid;
- measured bootstrap peak stability in a predefined tolerance;
- observed-versus-Ritz source consistency;
- held-out predictive behaviour with the frequency locked from training;
- calibrated false-positive rate under synthetic nulls;
- injection–recovery sensitivity.

**Not established here (requires independent evidence):**

- independent experimental replication or an independent dataset;
- a WCT physical mechanism or a universal atomic law;
- causal interpretation or NIST endorsement;
- significance beyond Monte Carlo resolution `1/(B+1)`;
- independence among all analysis variants (they share one line list).

```{r exec-claims}
show_tab(rd("final_claim_matrix.csv"))
```

# 2. Scope and claim boundaries

The scanner bins NIST line positions in `ell = ln(wavenumber / cm^-1)`, fits a
Poisson log-linear smooth baseline (`M0`), and tests for one additive
log-periodic mode `a cos(k ell) + b sin(k ell)` (`M1`). Here `k` is an
**angular** frequency, so the log-period is `2*pi/k` and the multiplicative
scale ratio is `exp(2*pi/k)`. `M0` is a statistical baseline, **not** a complete
physical atomic model.

# 3. Dataset accounting

Sequential, mutually exclusive dispositions reconcile exactly to the raw row
count; independent missing-value diagnostics are reported separately.

```{r flow}
show_tab(rd("dataset_flow_by_species.csv"))
```

```{r flow-fig}
#| out-width: 85%
p <- fig("fig01_dataset_flow.png"); if (!is.null(p)) knitr::include_graphics(p)
```

# 4. Canonical Fe II scan and 5. peak estimates / uncertainty

```{r peaks}
show_tab(rd("peak_estimates.csv"))
show_tab(rd("peak_confidence_intervals.csv"))
```

```{r boot-fig}
p <- fig("fig05_bootstrap_peak_distribution.png"); if (!is.null(p)) knitr::include_graphics(p)
```

# 6. Effect sizes

`z_null` is a **descriptive** standardized distance from the scan-global null
mean; the null is non-Gaussian, so it is not a Gaussian sigma. Raw `deltaD` is
not comparable across species without line and bin counts.

```{r eff}
show_tab(rd("effect_sizes.csv"))
```

```{r eff-fig}
p <- fig("fig07_effect_size_forest.png"); if (!is.null(p)) knitr::include_graphics(p)
p <- fig("fig06_null_ridgelines.png"); if (!is.null(p)) knitr::include_graphics(p)
```

# 7. Pointwise versus scan-global significance

Pointwise p is at the fixed prespecified `k`; scan-global p uses the maximum
statistic across the whole frequency scan (look-elsewhere corrected). Empirical
`p = (r+1)/(B+1)`; zero exceedances give the resolution floor `1/(B+1)`.

```{r sig}
show_tab(rd("significance_results.csv"))
```

# 8. Bin stability

```{r bins}
show_tab(rd("bin_stability_summary.csv"))
show_tab(rd("bin_grid_results.csv"))
```

```{r bins-fig}
p <- fig("fig03_frequency_by_bin_count.png"); if (!is.null(p)) knitr::include_graphics(p)
p <- fig("fig04_peak_stability.png"); if (!is.null(p)) knitr::include_graphics(p)
```

# 9. Model sensitivity

```{r spec}
show_tab(rd("specification_results.csv"), 20)
```

```{r spec-fig}
p <- fig("fig10_specification_curve.png"); if (!is.null(p)) knitr::include_graphics(p)
```

# 10. Neighbouring-ion comparison

See the effect-size table and ridgelines (section 6); neighbouring ions (Cr,
Mn, Co, Ni, Ti at ion II) are scanned with identical settings.

# 11. Model comparison

```{r mc}
show_tab(rd("model_comparison.csv"))
```

```{r mc-fig}
p <- fig("fig11_model_comparison.png"); if (!is.null(p)) knitr::include_graphics(p)
```

# 12. Observed-versus-Ritz comparison

```{r orr}
show_tab(rd("observed_ritz_replication.csv"))
```

```{r orr-fig}
p <- fig("fig08_observed_ritz_comparison.png"); if (!is.null(p)) knitr::include_graphics(p)
```

# 13. Blocked holdout validation

```{r ho}
show_tab(rd("holdout_results.csv"))
```

```{r ho-fig}
p <- fig("fig09_holdout_validation.png"); if (!is.null(p)) knitr::include_graphics(p)
```

# 14. Multiple-analysis correction

```{r mt}
show_tab(rd("multiple_testing.csv"))
```

```{r mt-fig}
p <- fig("fig12_multiplicity_comparison.png"); if (!is.null(p)) knitr::include_graphics(p)
```

# 15. Synthetic-null calibration

```{r cal}
show_tab(rd("null_calibration.csv"))
```

```{r cal-fig}
p <- fig("fig13_null_calibration.png"); if (!is.null(p)) knitr::include_graphics(p)
```

# 16. Injection–recovery power

```{r inj}
show_tab(rd("injection_recovery.csv"), 16)
```

```{r inj-fig}
p <- fig("fig14_injection_recovery_power.png"); if (!is.null(p)) knitr::include_graphics(p)
```

# 17. Python/R parity

```{r par}
show_tab(rd("python_r_parity.csv"))
```

```{r par-fig}
p <- fig("fig15_python_r_parity.png"); if (!is.null(p)) knitr::include_graphics(p)
p <- fig("fig16_evidence_dashboard.png"); if (!is.null(p)) knitr::include_graphics(p)
```

# 18. Limitations

- This is a computational reproduction; a different programming language is
  **not** independent replication.
- Observed and Ritz representations often describe the same transitions and are
  not independent datasets.
- The declared analyses share one NIST line list; multiplicity corrections are
  approximate and their conservativeness is discussed in section 14.
- Monte Carlo p-values cannot resolve below `1/(B+1)`.
- `M0` is a smooth statistical baseline, not a physical atomic model; no WCT
  mechanism, universal law, or NIST endorsement is claimed.

# 19. Reproducibility information

```{r repro}
cat("Analysis registry:", file.path("tables_r/statistical_audit", "analysis_registry.csv"), "\n")
reg <- rd("analysis_registry.csv")
if (!is.null(reg)) cat("Declared analyses:", nrow(reg),
                       "| seed:", unique(reg$seed)[1],
                       "| git:", unique(reg$git_commit)[1], "\n")
cat("Date:", format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z"), "\n\n")
cat("Exact development command:\n",
    "Rscript R/render_statistical_audit.R --fast --force --render-report true\n\n")
cat("Exact final command:\n",
    "Rscript R/render_statistical_audit.R --bootstrap-n 5000 --null-n 5000",
    "--calibration-n 10000 --injection-n 2000 --seed 20260517 --parallel true",
    "--force --strict --render-report true\n")
sessionInfo()
```
