---
title: "NIST Atomic-Line Log-Cosine Scan — Final Adversarial Statistical Audit"
subtitle: "Global calibration, frozen holdouts, alternative nulls, and preserved failures"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(fig.align = "center", out.width = "92%")
here_root <- normalizePath(file.path(dirname(knitr::current_input(dir = TRUE)), ".."),
                           mustWork = FALSE)
if (is.na(here_root) || !dir.exists(here_root)) here_root <- normalizePath("..", mustWork = FALSE)
TAB <- file.path(here_root, "tables_r/statistical_audit")
FIG <- file.path(here_root, "figures_r/statistical_audit")
rd <- function(f) {
  p <- file.path(TAB, f)
  if (file.exists(p)) utils::read.csv(p, stringsAsFactors = FALSE, check.names = FALSE) else NULL
}
fig <- function(f) {
  p <- file.path(FIG, f)
  if (file.exists(p)) p else NULL
}
show_tab <- function(df, n = NULL) {
  if (is.null(df)) return(cat("_table not available_"))
  if (!is.null(n)) df <- utils::head(df, n)
  knitr::kable(df)
}
```

# 1. Executive verdict

This report is the **final adversarial statistical audit** of the NIST atomic-line
log-cosine scan.  It is deliberately stricter than the development audit:
full-resolution Monte Carlo budgets are passed to every module without hidden
500/1,000-trial caps; the frequency is frozen before every confirmatory holdout;
alternative null generators are tested; and every failed, mixed, inconclusive,
or not-established claim is retained in a separate machine-readable table.

The central global-evidence table is shown below.  The column
`largest_reported_global_p` is the **largest** of the primary scan-global,
declared-family maximum-statistic, and worst alternative-null p-values.  It is a
conservative diagnostic summary only — **not** a newly combined formal p-value.

```{r global-evidence}
show_tab(rd("calibrated_global_evidence.csv"))
```

# 2. Claims — including failures

No result is promoted merely because another analysis passes.  In particular,
failed held-out predictive transfer, weak bin/specification stability, an
alternative-null failure, or a non-significant frozen holdout remains a failed or
mixed claim even if the in-sample peak is strong.

```{r final-claims}
show_tab(rd("final_adversarial_claim_matrix.csv"))
```

## 2.1 Non-passing claims retained explicitly

This table is mechanically generated from the claim matrix and contains every
verdict other than `pass`.

```{r failed-claims}
show_tab(rd("failed_claims.csv"))
```

# 3. Exact dataset and analysis provenance

```{r provenance}
show_tab(rd("final_adversarial_run_manifest.csv"))
show_tab(rd("final_adversarial_module_status.csv"))
show_tab(rd("analysis_registry.csv"), 30)
show_tab(rd("dataset_flow_by_species.csv"))
```

# 4. Primary peak and bootstrap uncertainty

The bootstrap uncertainty analysis resamples the observed dataset.  It measures
sampling/selection stability of the same line list; it is **not** independent
experimental replication.

```{r bootstrap}
show_tab(rd("peak_estimates.csv"))
show_tab(rd("peak_confidence_intervals.csv"))
show_tab(rd("peak_stability.csv"))
```

```{r bootstrap-figure}
p <- fig("fig05_bootstrap_peak_distribution.png")
if (!is.null(p)) knitr::include_graphics(p)
```

# 5. Full-scan look-elsewhere correction

Pointwise and scan-global evidence are kept separate.  The scan-global statistic
is the maximum deviance improvement over the full declared frequency grid.
Empirical p-values use `(r + 1)/(B + 1)`; zero exceedances are reported at the
Monte Carlo resolution floor rather than as p=0.

```{r global-scan}
show_tab(rd("significance_results.csv"))
```

# 6. Declared-family multiplicity correction

The audit reports BH-FDR, Holm, Bonferroni, and a family maximum-statistic
calibration across the declared analysis family.  Shared underlying line lists
make the family dependence nontrivial; this dependence is preserved as a stated
limitation rather than hidden.

```{r multiplicity}
show_tab(rd("multiple_testing.csv"))
```

```{r multiplicity-figure}
p <- fig("fig12_multiplicity_comparison.png")
if (!is.null(p)) knitr::include_graphics(p)
```

# 7. Adversarial alternative null models

The primary canonical null is supplemented with four stress tests:

1. fitted Poisson with the smoothing baseline re-estimated per replicate;
2. conditional multinomial counts with fixed total line count;
3. a negative-binomial overdispersion null;
4. a block-permuted Pearson-residual diagnostic that preserves local residual
   chunks while destroying coherent long-range ordering.

The conservative result is the **largest** scan-global p-value among the declared
nulls.  The block-residual model is explicitly a misspecification diagnostic,
not asserted to be the unique physical null.

```{r alt-nulls}
show_tab(rd("alternative_null_results.csv"))
```

```{r alt-nulls-figure}
p <- fig("fig17_alternative_nulls.png")
if (!is.null(p)) knitr::include_graphics(p)
```

# 8. Synthetic-null calibration

Synthetic-null calibration asks whether the scan-global decision rule achieves
approximately its nominal type-I error rate under the declared canonical null.
The interval is reported because finite Monte Carlo uncertainty can be material.

```{r calibration}
show_tab(rd("null_calibration.csv"))
```

```{r calibration-figure}
p <- fig("fig13_null_calibration.png")
if (!is.null(p)) knitr::include_graphics(p)
```

# 9. Frozen blocked holdouts

For every confirmatory holdout, `k` is selected **only on training bins and then
locked**.  The test block is evaluated at that fixed frequency.  Test-block
rescans, where present, remain exploratory and cannot rescue a failed locked-k
result.

```{r holdout}
show_tab(rd("holdout_results.csv"))
```

```{r holdout-figure}
p <- fig("fig09_holdout_validation.png")
if (!is.null(p)) knitr::include_graphics(p)
```

# 10. Preprocessing and specification sensitivity

The predefined multiverse varies baseline smoothing, polynomial degree, bin
count, and line-position representation.  A strong canonical result is not
reported as globally robust if a substantial part of this declared multiverse
selects a different frequency region.

```{r sensitivity}
show_tab(rd("bin_stability_summary.csv"))
show_tab(rd("bin_grid_results.csv"))
show_tab(rd("specification_results.csv"))
show_tab(rd("observed_ritz_replication.csv"))
```

```{r sensitivity-figures}
p <- fig("fig10_specification_curve.png"); if (!is.null(p)) knitr::include_graphics(p)
p <- fig("specification_heatmap.png"); if (!is.null(p)) knitr::include_graphics(p)
```

# 11. In-sample versus held-out model comparison

AIC/BIC improvement after selecting a frequency on the same data does not by
itself demonstrate predictive transfer.  The held-out log-likelihood comparison
is therefore reported separately and any negative transfer is retained as an
explicit failed claim.

```{r model-comparison}
show_tab(rd("model_comparison.csv"))
```

```{r model-comparison-figure}
p <- fig("fig11_model_comparison.png")
if (!is.null(p)) knitr::include_graphics(p)
```

# 12. Injection and recovery

Injection/recovery reports both detection of any significant peak and recovery
inside the correct predefined frequency region.  These are not interchangeable:
a detector that fires at the wrong frequency has not recovered the injected
structure.

```{r injection}
show_tab(rd("injection_recovery.csv"))
```

```{r injection-figure}
p <- fig("fig14_injection_recovery_power.png")
if (!is.null(p)) knitr::include_graphics(p)
```

# 13. Computational parity

```{r parity}
show_tab(rd("python_r_parity.csv"))
```

```{r parity-figure}
p <- fig("fig15_python_r_parity.png")
if (!is.null(p)) knitr::include_graphics(p)
```

# 14. Interpretation boundaries

This audit can support or weaken a claim of **statistical structure in this
analysis pipeline**.  It cannot by itself establish:

- independent experimental confirmation;
- independence of observed/Ritz representations or blocked subsets drawn from
  the same transition list;
- a unique physical null model;
- a WCT mechanism, causal explanation, or universal atomic law;
- NIST endorsement of the analysis.

The report therefore treats computational parity, same-dataset bootstrap
stability, frozen holdouts, alternative-null robustness, and external physical
replication as different evidentiary categories.

# 15. Reproduction command

The strict final runner does not use the development caps in
`render_statistical_audit.R`.

```{r reproduce}
cat("Rscript R/run_final_adversarial_audit.R --parallel true --render-report true\n\n")
manifest <- rd("final_adversarial_run_manifest.csv")
if (!is.null(manifest)) show_tab(manifest)
cat("\nSession information for this render:\n")
sessionInfo()
```
